---
title: "EDS241: Assignment 4"
author: "Shale Hunter"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "readr")

for (i in packages) {
    require(i,character.only=TRUE)
  }


options(scipen=10) 


```

## Import & Clean Data

```{r}
df = read_csv("EU_sardines.csv")
```

```{r}
df = df %>% 
  mutate(log_price = log(price_euro_kg),
         log_volume = log(volume_sold_kg))
```

# Questions

### (a) Estimate a bivariate regression of log(volume_sold_kg) on log(price euro_kg). What is the price elasticity of demand for sardines? Test the null hypothesis that the price elasticity is equal to -1.

```{r}
m1 = estimatr::lm_robust(data = df, log_price ~ log_volume)
summary(m1)
car::linearHypothesis(m1, "log_volume=-1")
```

**The model above gives a price elasticity of -0.06, which a linear hypothesis test shows rejects the null hypothesis of a price elasticity of -1 at p < 0.0001.**

### (b) Like in Lecture 8 (see the IV.R script), we will use wind_m_s as an instrument for log(price_euro_kg). To begin, estimate the first-stage regression relating log(price_euro_kg) to wind_m_s. Interpret the estimated coefficient on wind speed. Does it have the expected sign? Also test for the relevance of the instrument and whether it is a “weak” instrument by reporting the proper F-statistic.

```{r}
mw = estimatr::lm_robust(data = df, log_price ~ wind_m_s)
summary(mw)
```

**The `wind_m_s` coefficient of 0.67 means that a 1 m/s increase in wind speed should correspond to a 0.067 euro/kg increase in price (both values logged). It makes sense that this relationship is positive because higher winds means it is harder/more dangerous to catch fish (and supply may therefore go down), so cost should increase if demand is unaffected. With an F-statistic of 144.7, wind speed is a non-weak relevant instrument.**

### (c) Estimate the TSLS estimator of the price elasticity of demand for sardines using wind_m_s as an instrument for log(price_euro_kg). What is the estimated price elasticity of demand for sardines?

```{r}
tsls = AER::ivreg(data = df, log_price ~ log_volume | wind_m_s)

summary(tsls)
```

**The estimated price elasticity of demand for sardines is -0.92.**

### (d) Repeat the exercise in (c), but include fixed effects for each year, month, and country. [Hint: you can use the command “as.factor(country) + as.factor(year) +as.factor(month)” to the ivreg function in R]. Report the estimated price elasticity of demand and the F-statistic testing for relevant and non-weak instruments.

```{r}
tslsf = AER::ivreg(data = df, log_price ~ log_volume | wind_m_s + as.factor(country) + as.factor(year) + as.factor(month))

summary(tslsf)
```

**Controling for fixed effects of country, year, and month, the new price elasticity of demand is 0.0226. The F-statistic of **

